---
title: "Analysis Results"
author: "Linh Nguyen"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: 'hide'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# I. Meta-data

```{r library, message = F}
require(tidyverse)
require(rio)
require(mice)
require(cowplot)
require(reshape2)
require(lavaan)
require(semPlot)
require(kableExtra)
require(summarytools)
options(digits = 3)
set.seed(202104)
print(sessionInfo(), locale = FALSE)
```

# II. MPQ cleaning and descriptives

```{r import-data}
#load data
mpq <- rio::import(file = "./Data/SIBS_LONGITUDINAL_PERSONALITY_20210413.csv") %>% 
  select(-1)

#aggression items
mpq$AG_IN <- mpq %>% 
  select(starts_with("AG") & ends_with("IN")) %>% 
  rowMeans(na.rm = T)
mpq$AG_FU1 <- mpq %>% 
  select(starts_with("AG") & ends_with("FU1")) %>% 
  rowMeans(na.rm = T)
mpq$AG_FU2 <- mpq %>% 
  select(starts_with("AG") & ends_with("FU2")) %>% 
  rowMeans(na.rm = T)
mpq$AG_FU3 <- mpq %>% 
  select(starts_with("AG") & ends_with("FU3")) %>% 
  rowMeans(na.rm = T)

#control items
mpq$CON_IN <- mpq %>% 
  select(starts_with("CON") & ends_with("IN")) %>% 
  rowMeans(na.rm = T)
mpq$CON_FU1 <- mpq %>% 
  select(starts_with("CON") & ends_with("FU1")) %>% 
  rowMeans(na.rm = T)
mpq$CON_FU2 <- mpq %>% 
  select(starts_with("CON") & ends_with("FU2")) %>% 
  rowMeans(na.rm = T)
mpq$CON_FU3 <- mpq %>% 
  select(starts_with("CON") & ends_with("FU3")) %>% 
  rowMeans(na.rm = T)

#harm avoidance items
mpq$HA_IN <- mpq %>% 
  select(starts_with("HA") & ends_with("IN")) %>% 
  rowMeans(na.rm = T)
mpq$HA_FU1 <- mpq %>% 
  select(starts_with("HA") & ends_with("FU1")) %>% 
  rowMeans(na.rm = T)
mpq$HA_FU2 <- mpq %>% 
  select(starts_with("HA") & ends_with("FU2")) %>% 
  rowMeans(na.rm = T)
mpq$HA_FU3 <- mpq %>% 
  select(starts_with("HA") & ends_with("FU3")) %>% 
  rowMeans(na.rm = T)

#clean up NaN
mpq$AG_IN[is.nan(mpq$AG_IN)]<-NA
mpq$AG_FU1[is.nan(mpq$AG_FU1)]<-NA
mpq$AG_FU2[is.nan(mpq$AG_FU2)]<-NA
mpq$AG_FU3[is.nan(mpq$AG_FU3)]<-NA
mpq$CON_IN[is.nan(mpq$CON_IN)]<-NA
mpq$CON_FU1[is.nan(mpq$CON_FU1)]<-NA
mpq$CON_FU2[is.nan(mpq$CON_FU2)]<-NA
mpq$CON_FU3[is.nan(mpq$CON_FU3)]<-NA
mpq$HA_IN[is.nan(mpq$HA_IN)]<-NA
mpq$HA_FU1[is.nan(mpq$HA_FU1)]<-NA
mpq$HA_FU2[is.nan(mpq$HA_FU2)]<-NA
mpq$HA_FU3[is.nan(mpq$HA_FU3)]<-NA

#create wide dyad dataframe
mpq <- mpq[order(mpq$ID),]
mpq$P <- rep(c(1:2), length.out = nrow(mpq)) #person dummy var
yo <- mpq %>% select(IDYRFAM, BDAY, P) %>% 
  mutate(BDAY = as.character(BDAY),
         P = as.character(P))
yo <- reshape2::dcast(yo, IDYRFAM ~ P, value.var = "BDAY") #wide data with 2 siblings per row
yo <- yo %>% mutate(`1` = as.Date(`1`),
                    `2` = as.Date(`2`))
yo <- yo %>% mutate(older = ifelse(`1`<`2`, "P1", "P2")) %>% 
  select(IDYRFAM, older)
mpq <- merge(mpq, yo)
mpq <- mpq %>% 
  mutate(yo = ifelse(P == 1 & older == "P2" | P == 2 & older == "P1", "y", "o")) %>%  #y = younger, o = older
  select(-P, -older)

young <- mpq %>%      #data with only younger sib
  filter(yo == "y")
names(young) <- paste0(names(young), "_y", sep = "")
young <- young %>% rename(IDYRFAM = IDYRFAM_y)

old <- mpq %>%        #data with only older sib
  filter(yo == "o")
names(old) <- paste0(names(old), "_o", sep = "")
old <- old %>% rename(IDYRFAM = IDYRFAM_o)

yo <- merge(young, old)
rm(young, old)

#create adopted/bio variable
yo <- yo %>% 
  mutate(adopt = ifelse(IDAB_y == 1 | IDAB_o == 1, 1, 0))
```

### Demographics

```{r demo, results = "asis"}
#descriptives
mpq %>% 
  select(IDSEX:FU3_AGE) %>% 
  descr(stats = "common", order = "p")
```

The average time between assessments is:

- Between intake and follow-up 1: `r as.numeric(mean(as.Date(mpq$NCDATE_FU1) - as.Date(mpq$NCDATE_IN), na.rm = T))/365` years.
- Between follow-up 1 and follow-up 2: `r as.numeric(mean(as.Date(mpq$NCDATE_FU2) - as.Date(mpq$NCDATE_FU1), na.rm = T))/365` years.
- Between follow-up 3 and follow-up 2: `r as.numeric(mean(as.Date(mpq$NCDATE_FU3) - as.Date(mpq$NCDATE_FU2), na.rm = T))/365` years.

### Scale scores

```{r scores, message = F, warning = F, results = "asis"}
#descriptives
mpq %>% select(AG_IN:HA_FU3) %>% descr(stats = "common", order = "p")

#random sample for plots
sample = sample(mpq$ID, size = 100)
sample <- mpq %>% filter(ID %in% c(sample))
sample <- reshape(sample, direction = "long",
                varying = list(c("IN_AGE", "FU1_AGE", "FU2_AGE", "FU3_AGE"),
                               c("AG_IN", "AG_FU1", "AG_FU2", "AG_FU3"),
                               c("CON_IN", "CON_FU1", "CON_FU2", "CON_FU3"),
                               c("HA_IN", "HA_FU1", "HA_FU2", "HA_FU3")),
                timevar = "time",
                times = c(0,1,2,3),
                v.names = c("age","AG","CON","HA"),
                idvar = c("ID"))
row.names(sample) <- 1:nrow(sample)

#plot by timepoints
pAG <- ggplot(data = sample,
               aes(x = time, y = AG, group = ID)) + 
    geom_line (linetype = "dashed")+ 
    geom_point(size = 0.5) +
    theme_bw () +
    stat_summary(aes(data=sample$AG,group=1),fun=mean,geom="line",lwd = 1.5, color= "red")
pCON <- ggplot(data = sample,
                aes(x = time, y = CON, group = ID)) + 
    geom_line (linetype = "dashed")+ 
    geom_point(size = 0.5) +
    theme_bw () +
    stat_summary(aes(data=sample$CON,group=1),fun=mean,geom="line",lwd = 1.5, color= "red")
pHA <- ggplot(data = sample,
               aes(x = time, y = HA, group = ID)) + 
    geom_line (linetype = "dashed")+ 
    geom_point(size = 0.5) +
    theme_bw () +
    stat_summary(aes(data=sample$HA,group=1),fun=mean,geom="line",lwd = 1.5, color= "red")

#plot by age
paAG <- ggplot(data = sample,
               aes(x = age, y = AG, group = ID)) + 
    geom_line (linetype = "dashed")+ 
    geom_point(size = 0.25) + 
    theme_bw () +
    stat_smooth(aes(data=sample$AG,group=1),method="lm",formula=y ~ poly(x, 2),lwd = 1.5, color= "red")
paCON <- ggplot(data = sample,
                aes(x = age, y = CON, group = ID)) + 
    geom_line (linetype = "dashed")+ 
    geom_point(size = 0.25) +
    theme_bw () +
    stat_smooth(aes(data=sample$CON,group=1),method="lm",formula=y ~ poly(x, 2),lwd = 1.5, color= "red")
paHA <- ggplot(data = sample,
               aes(x = age, y = HA, group = ID)) + 
    geom_line (linetype = "dashed")+ 
    geom_point(size = 0.25) +
    theme_bw () +
    stat_smooth(aes(data=sample$HA,group=1),method="lm",formula=y ~ poly(x, 2),lwd = 1.5, color= "red")
  
cowplot::plot_grid(pAG, paAG,
                   pCON, paCON, 
                   pHA, paHA,
                   nrow = 3, ncol = 2)

rm(sample, pAG, paAG, pCON, paCON, pHA, paHA)
```

There are `r 1231 - sum(is.na(mpq$AG_IN))` participants with complete responses at intake, `r 1231 - sum(is.na(mpq$AG_FU1))` at follow-up 1, `r 1231 - sum(is.na(mpq$AG_FU2))`  at follow-up 2, and `r 1231 - sum(is.na(mpq$AG_FU3))` at follow-up 3.

There are `r length(unique(mpq[!is.na(mpq$AG_IN),]$IDYRFAM))` families with at least 1 sibling at intake, `r length(unique(mpq[!is.na(mpq$AG_FU1),]$IDYRFAM))` at follow-up 1, `r length(unique(mpq[!is.na(mpq$AG_FU2),]$IDYRFAM))` at follow-up 2, and `r length(unique(mpq[!is.na(mpq$AG_FU3),]$IDYRFAM))` at follow-up 3.

There are `r sum(duplicated(mpq[!is.na(mpq$AG_IN),]$IDYRFAM))` families with both siblings at intake, `r sum(duplicated(mpq[!is.na(mpq$AG_FU1),]$IDYRFAM))` at follow-up 1, `r sum(duplicated(mpq[!is.na(mpq$AG_FU2),]$IDYRFAM))` at follow-up 2, and `r sum(duplicated(mpq[!is.na(mpq$AG_FU3),]$IDYRFAM))` at follow-up 3.


# III. Analyses

## Random parcel sampling

```{r parcel}
par <- c(1:18)
par1 <- sample(par, size = 6) #parcel 1
par2 <- sample(par[which(!par %in% par1)], size = 6) #parcel 2
par3 <- par[which(!par %in% par1 & !par %in% par2)] #parcel 3

# aggression AG 
yo <- yo %>% 
  mutate(# older sibling
         AG_IN_o_p1 = rowMeans(select(yo, AG12_IN_o, AG18_IN_o, AG4_IN_o, AG15_IN_o, AG6_IN_o, AG2_IN_o), na.rm = T),
         AG_IN_o_p2 = rowMeans(select(yo, AG10_IN_o, AG17_IN_o, AG7_IN_o, AG13_IN_o, AG1_IN_o, AG3_IN_o), na.rm = T),
         AG_IN_o_p3 = rowMeans(select(yo, AG5_IN_o, AG8_IN_o, AG9_IN_o, AG11_IN_o, AG14_IN_o, AG16_IN_o), na.rm = T),
         
         AG_FU1_o_p1 = rowMeans(select(yo, AG12_FU1_o, AG18_FU1_o, AG4_FU1_o, AG15_FU1_o, AG6_FU1_o, AG2_FU1_o), na.rm = T),
         AG_FU1_o_p2 = rowMeans(select(yo, AG10_FU1_o, AG17_FU1_o, AG7_FU1_o, AG13_FU1_o, AG1_FU1_o, AG3_FU1_o), na.rm = T),
         AG_FU1_o_p3 = rowMeans(select(yo, AG5_FU1_o, AG8_FU1_o, AG9_FU1_o, AG11_FU1_o, AG14_FU1_o, AG16_FU1_o), na.rm = T),
         
         AG_FU2_o_p1 = rowMeans(select(yo, AG12_FU2_o, AG18_FU2_o, AG4_FU2_o, AG15_FU2_o, AG6_FU2_o, AG2_FU2_o), na.rm = T),
         AG_FU2_o_p2 = rowMeans(select(yo, AG10_FU2_o, AG17_FU2_o, AG7_FU2_o, AG13_FU2_o, AG1_FU2_o, AG3_FU2_o), na.rm = T),
         AG_FU2_o_p3 = rowMeans(select(yo, AG5_FU2_o, AG8_FU2_o, AG9_FU2_o, AG11_FU2_o, AG14_FU2_o, AG16_FU2_o), na.rm = T),
         
         AG_FU3_o_p1 = rowMeans(select(yo, AG12_FU3_o, AG18_FU3_o, AG4_FU3_o, AG6_FU3_o, AG2_FU3_o), na.rm = T),
         AG_FU3_o_p2 = rowMeans(select(yo, AG10_FU3_o, AG7_FU3_o), na.rm = T),
         AG_FU3_o_p3 = rowMeans(select(yo, AG5_FU3_o, AG8_FU3_o, AG11_FU3_o, AG14_FU3_o, AG16_FU3_o), na.rm = T),
         
         # younger sibling
         AG_IN_y_p1 = rowMeans(select(yo, AG12_IN_y, AG18_IN_y, AG4_IN_y, AG15_IN_y, AG6_IN_y, AG2_IN_y), na.rm = T),
         AG_IN_y_p2 = rowMeans(select(yo, AG10_IN_y, AG17_IN_y, AG7_IN_y, AG13_IN_y, AG1_IN_y, AG3_IN_y), na.rm = T),
         AG_IN_y_p3 = rowMeans(select(yo, AG5_IN_y, AG8_IN_y, AG9_IN_y, AG11_IN_y, AG14_IN_y, AG16_IN_y), na.rm = T),
         
         AG_FU1_y_p1 = rowMeans(select(yo, AG12_FU1_y, AG18_FU1_y, AG4_FU1_y, AG15_FU1_y, AG6_FU1_y, AG2_FU1_y), na.rm = T),
         AG_FU1_y_p2 = rowMeans(select(yo, AG10_FU1_y, AG17_FU1_y, AG7_FU1_y, AG13_FU1_y, AG1_FU1_y, AG3_FU1_y), na.rm = T),
         AG_FU1_y_p3 = rowMeans(select(yo, AG5_FU1_y, AG8_FU1_y, AG9_FU1_y, AG11_FU1_y, AG14_FU1_y, AG16_FU1_y), na.rm = T),
         
         AG_FU2_y_p1 = rowMeans(select(yo, AG12_FU2_y, AG18_FU2_y, AG4_FU2_y, AG15_FU2_y, AG6_FU2_y, AG2_FU2_y), na.rm = T),
         AG_FU2_y_p2 = rowMeans(select(yo, AG10_FU2_y, AG17_FU2_y, AG7_FU2_y, AG13_FU2_y, AG1_FU2_y, AG3_FU2_y), na.rm = T),
         AG_FU2_y_p3 = rowMeans(select(yo, AG5_FU2_y, AG8_FU2_y, AG9_FU2_y, AG11_FU2_y, AG14_FU2_y, AG16_FU2_y), na.rm = T),
         
         AG_FU3_y_p1 = rowMeans(select(yo, AG12_FU3_y, AG18_FU3_y, AG4_FU3_y, AG6_FU3_y, AG2_FU3_y), na.rm = T),
         AG_FU3_y_p2 = rowMeans(select(yo, AG10_FU3_y, AG7_FU3_y), na.rm = T),
         AG_FU3_y_p3 = rowMeans(select(yo, AG5_FU3_y, AG8_FU3_y, AG11_FU3_y, AG14_FU3_y, AG16_FU3_y), na.rm = T))

# control CON 
yo <- yo %>% 
  mutate(# older sibling
         CON_IN_o_p1 = rowMeans(select(yo, CON12_IN_o, CON18_IN_o, CON4_IN_o, CON15_IN_o, CON6_IN_o, CON2_IN_o), na.rm = T),
         CON_IN_o_p2 = rowMeans(select(yo, CON10_IN_o, CON17_IN_o, CON7_IN_o, CON13_IN_o, CON1_IN_o, CON3_IN_o), na.rm = T),
         CON_IN_o_p3 = rowMeans(select(yo, CON5_IN_o, CON8_IN_o, CON9_IN_o, CON11_IN_o, CON14_IN_o, CON16_IN_o), na.rm = T),
         
         CON_FU1_o_p1 = rowMeans(select(yo, CON12_FU1_o, CON18_FU1_o, CON4_FU1_o, CON15_FU1_o, CON6_FU1_o, CON2_FU1_o), na.rm = T),
         CON_FU1_o_p2 = rowMeans(select(yo, CON10_FU1_o, CON17_FU1_o, CON7_FU1_o, CON13_FU1_o, CON1_FU1_o, CON3_FU1_o), na.rm = T),
         CON_FU1_o_p3 = rowMeans(select(yo, CON5_FU1_o, CON8_FU1_o, CON9_FU1_o, CON11_FU1_o, CON14_FU1_o, CON16_FU1_o), na.rm = T),
         
         CON_FU2_o_p1 = rowMeans(select(yo, CON12_FU2_o, CON18_FU2_o, CON4_FU2_o, CON15_FU2_o, CON6_FU2_o, CON2_FU2_o), na.rm = T),
         CON_FU2_o_p2 = rowMeans(select(yo, CON10_FU2_o, CON17_FU2_o, CON7_FU2_o, CON13_FU2_o, CON1_FU2_o, CON3_FU2_o), na.rm = T),
         CON_FU2_o_p3 = rowMeans(select(yo, CON5_FU2_o, CON8_FU2_o, CON9_FU2_o, CON11_FU2_o, CON14_FU2_o, CON16_FU2_o), na.rm = T),
         
         CON_FU3_o_p1 = rowMeans(select(yo, CON4_FU3_o, CON15_FU3_o, CON6_FU3_o, CON2_FU3_o), na.rm = T),
         CON_FU3_o_p2 = rowMeans(select(yo, CON10_FU3_o, CON17_FU3_o, CON7_FU3_o), na.rm = T),
         CON_FU3_o_p3 = rowMeans(select(yo, CON5_FU3_o, CON8_FU3_o, CON9_FU3_o, CON16_FU3_o), na.rm = T),
         
         # younger sibling
         CON_IN_y_p1 = rowMeans(select(yo, CON12_IN_y, CON18_IN_y, CON4_IN_y, CON15_IN_y, CON6_IN_y, CON2_IN_y), na.rm = T),
         CON_IN_y_p2 = rowMeans(select(yo, CON10_IN_y, CON17_IN_y, CON7_IN_y, CON13_IN_y, CON1_IN_y, CON3_IN_y), na.rm = T),
         CON_IN_y_p3 = rowMeans(select(yo, CON5_IN_y, CON8_IN_y, CON9_IN_y, CON11_IN_y, CON14_IN_y, CON16_IN_y), na.rm = T),
         
         CON_FU1_y_p1 = rowMeans(select(yo, CON12_FU1_y, CON18_FU1_y, CON4_FU1_y, CON15_FU1_y, CON6_FU1_y, CON2_FU1_y), na.rm = T),
         CON_FU1_y_p2 = rowMeans(select(yo, CON10_FU1_y, CON17_FU1_y, CON7_FU1_y, CON13_FU1_y, CON1_FU1_y, CON3_FU1_y), na.rm = T),
         CON_FU1_y_p3 = rowMeans(select(yo, CON5_FU1_y, CON8_FU1_y, CON9_FU1_y, CON11_FU1_y, CON14_FU1_y, CON16_FU1_y), na.rm = T),
         
         CON_FU2_y_p1 = rowMeans(select(yo, CON12_FU2_y, CON18_FU2_y, CON4_FU2_y, CON15_FU2_y, CON6_FU2_y, CON2_FU2_y), na.rm = T),
         CON_FU2_y_p2 = rowMeans(select(yo, CON10_FU2_y, CON17_FU2_y, CON7_FU2_y, CON13_FU2_y, CON1_FU2_y, CON3_FU2_y), na.rm = T),
         CON_FU2_y_p3 = rowMeans(select(yo, CON5_FU2_y, CON8_FU2_y, CON9_FU2_y, CON11_FU2_y, CON14_FU2_y, CON16_FU2_y), na.rm = T),
         
         CON_FU3_y_p1 = rowMeans(select(yo, CON4_FU3_y, CON15_FU3_y, CON6_FU3_y, CON2_FU3_y), na.rm = T),
         CON_FU3_y_p2 = rowMeans(select(yo, CON10_FU3_y, CON17_FU3_y, CON7_FU3_y), na.rm = T),
         CON_FU3_y_p3 = rowMeans(select(yo, CON5_FU3_y, CON8_FU3_y, CON9_FU3_y, CON16_FU3_y), na.rm = T))

# harm avoidance HA
yo <- yo %>% 
    mutate(# older sibling
         HA_IN_o_p1 = rowMeans(select(yo, HA12_IN_o, HA18_IN_o, HA4_IN_o, HA15_IN_o, HA6_IN_o, HA2_IN_o), na.rm = T),
         HA_IN_o_p2 = rowMeans(select(yo, HA10_IN_o, HA17_IN_o, HA7_IN_o, HA13_IN_o, HA1_IN_o, HA3_IN_o), na.rm = T),
         HA_IN_o_p3 = rowMeans(select(yo, HA5_IN_o, HA8_IN_o, HA9_IN_o, HA11_IN_o, HA14_IN_o, HA16_IN_o), na.rm = T),
         
         HA_FU1_o_p1 = rowMeans(select(yo, HA12_FU1_o, HA18_FU1_o, HA4_FU1_o, HA15_FU1_o, HA6_FU1_o, HA2_FU1_o), na.rm = T),
         HA_FU1_o_p2 = rowMeans(select(yo, HA10_FU1_o, HA17_FU1_o, HA7_FU1_o, HA13_FU1_o, HA1_FU1_o, HA3_FU1_o), na.rm = T),
         HA_FU1_o_p3 = rowMeans(select(yo, HA5_FU1_o, HA8_FU1_o, HA9_FU1_o, HA11_FU1_o, HA14_FU1_o, HA16_FU1_o), na.rm = T),
         
         HA_FU2_o_p1 = rowMeans(select(yo, HA12_FU2_o, HA18_FU2_o, HA4_FU2_o, HA15_FU2_o, HA6_FU2_o, HA2_FU2_o), na.rm = T),
         HA_FU2_o_p2 = rowMeans(select(yo, HA10_FU2_o, HA17_FU2_o, HA7_FU2_o, HA13_FU2_o, HA1_FU2_o, HA3_FU2_o), na.rm = T),
         HA_FU2_o_p3 = rowMeans(select(yo, HA5_FU2_o, HA8_FU2_o, HA9_FU2_o, HA11_FU2_o, HA14_FU2_o, HA16_FU2_o), na.rm = T),
         
         HA_FU3_o_p1 = rowMeans(select(yo, HA18_FU3_o, HA4_FU3_o, HA6_FU3_o, HA2_FU3_o), na.rm = T),
         HA_FU3_o_p2 = rowMeans(select(yo, HA10_FU3_o, HA7_FU3_o, HA13_FU3_o, HA1_FU3_o, HA3_FU3_o), na.rm = T),
         HA_FU3_o_p3 = rowMeans(select(yo, HA8_FU3_o, HA9_FU3_o, HA16_FU3_o), na.rm = T),
         
         # younger sibling
         HA_IN_y_p1 = rowMeans(select(yo, HA12_IN_y, HA18_IN_y, HA4_IN_y, HA15_IN_y, HA6_IN_y, HA2_IN_y), na.rm = T),
         HA_IN_y_p2 = rowMeans(select(yo, HA10_IN_y, HA17_IN_y, HA7_IN_y, HA13_IN_y, HA1_IN_y, HA3_IN_y), na.rm = T),
         HA_IN_y_p3 = rowMeans(select(yo, HA5_IN_y, HA8_IN_y, HA9_IN_y, HA11_IN_y, HA14_IN_y, HA16_IN_y), na.rm = T),
         
         HA_FU1_y_p1 = rowMeans(select(yo, HA12_FU1_y, HA18_FU1_y, HA4_FU1_y, HA15_FU1_y, HA6_FU1_y, HA2_FU1_y), na.rm = T),
         HA_FU1_y_p2 = rowMeans(select(yo, HA10_FU1_y, HA17_FU1_y, HA7_FU1_y, HA13_FU1_y, HA1_FU1_y, HA3_FU1_y), na.rm = T),
         HA_FU1_y_p3 = rowMeans(select(yo, HA5_FU1_y, HA8_FU1_y, HA9_FU1_y, HA11_FU1_y, HA14_FU1_y, HA16_FU1_y), na.rm = T),
         
         HA_FU2_y_p1 = rowMeans(select(yo, HA12_FU2_y, HA18_FU2_y, HA4_FU2_y, HA15_FU2_y, HA6_FU2_y, HA2_FU2_y), na.rm = T),
         HA_FU2_y_p2 = rowMeans(select(yo, HA10_FU2_y, HA17_FU2_y, HA7_FU2_y, HA13_FU2_y, HA1_FU2_y, HA3_FU2_y), na.rm = T),
         HA_FU2_y_p3 = rowMeans(select(yo, HA5_FU2_y, HA8_FU2_y, HA9_FU2_y, HA11_FU2_y, HA14_FU2_y, HA16_FU2_y), na.rm = T),
         
         HA_FU3_y_p1 = rowMeans(select(yo, HA18_FU3_y, HA4_FU3_y, HA6_FU3_y, HA2_FU3_y), na.rm = T),
         HA_FU3_y_p2 = rowMeans(select(yo, HA10_FU3_y, HA7_FU3_y, HA13_FU3_y, HA1_FU3_y, HA3_FU3_y), na.rm = T),
         HA_FU3_y_p3 = rowMeans(select(yo, HA8_FU3_y, HA9_FU3_y, HA16_FU3_y), na.rm = T))

  
# clean up NaN
for (row in seq(nrow(yo))) {
  for (col in seq(ncol(yo))) {
    if (is.nan(yo[row, col])) {
      yo[row, col] <- NA
    }}} # END for row/col LOOP
```

For SEM analyses of each personality variable, the three parcels were randomized to contain items numbered:

- Parcel 1: `r paste0(sort(par1))`
- Parcel 2: `r paste0(sort(par2))`
- Parcel 3: `r paste0(sort(par3))`


## Adopted vs. Non-adopted sibling pairs

### Aggression AG

```{r adopt-AG, warning = F, cache = T}
sibAG <- '
#younger sibling
AG0y =~   AG_IN_y_p1  + p2 * AG_IN_y_p2   + p3 * AG_IN_y_p3
AG3y =~   AG_FU1_y_p1 + p2 * AG_FU1_y_p2  + p3 * AG_FU1_y_p3
AG7y =~   AG_FU2_y_p1 + p2 * AG_FU2_y_p2  + p3 * AG_FU2_y_p3
AG16y =~  AG_FU3_y_p1 + p2 * AG_FU3_y_p2  + p3 * AG_FU3_y_p3
                   
#older sibling
AG0o =~   AG_IN_o_p1  + p2 * AG_IN_o_p2   + p3 * AG_IN_o_p3
AG3o =~   AG_FU1_o_p1 + p2 * AG_FU1_o_p2  + p3 * AG_FU1_o_p3
AG7o =~   AG_FU2_o_p1 + p2 * AG_FU2_o_p2  + p3 * AG_FU2_o_p3
AG16o =~  AG_FU3_o_p1 + p2 * AG_FU3_o_p2  + p3 * AG_FU3_o_p3

#latent intercept and slopes
iy =~ 1*AG0y + 1*AG3y + 1*AG7y + 1*AG16y
io =~ 1*AG0o + 1*AG3o + 1*AG7o + 1*AG16o
sy =~ 0*AG0y + 3.3*AG3y + 7.6*AG7y + 17*AG16y
so =~ 0*AG0o + 3.3*AG3o + 7.6*AG7o + 17*AG16o
iy ~ 1
io ~ 1
sy ~ 1
so ~ 1

#fix zero latent intercepts
AG0y ~ 0*1
AG3y ~ 0*1
AG7y ~ 0*1
AG16y ~ 0*1
AG0o ~ 0*1
AG3o ~ 0*1
AG7o ~ 0*1
AG16o ~ 0*1

#fix zero/equal item intercepts
AG_IN_y_p1  ~ 0*1
AG_FU1_y_p1 ~ 0*1
AG_FU2_y_p1 ~ 0*1
AG_FU3_y_p1 ~ 0*1
AG_IN_o_p1  ~ 0*1
AG_FU1_o_p1 ~ 0*1
AG_FU2_o_p1 ~ 0*1
AG_FU3_o_p1 ~ 0*1

AG_IN_y_p2  ~ par2*1
AG_FU1_y_p2 ~ par2*1
AG_FU2_y_p2 ~ par2*1
AG_FU3_y_p2 ~ par2*1
AG_IN_o_p2  ~ par2*1
AG_FU1_o_p2 ~ par2*1
AG_FU2_o_p2 ~ par2*1
AG_FU3_o_p2 ~ par2*1

AG_IN_y_p3  ~ par3*1
AG_FU1_y_p3 ~ par3*1
AG_FU2_y_p3 ~ par3*1
AG_FU3_y_p3 ~ par3*1
AG_IN_o_p3  ~ par3*1
AG_FU1_o_p3 ~ par3*1
AG_FU2_o_p3 ~ par3*1
AG_FU3_o_p3 ~ par3*1

#residual covariances
AG_IN_y_p1 ~~ c(cov11)*AG_FU1_y_p1 + c(cov12)*AG_FU2_y_p1 + c(cov13)*AG_FU3_y_p1 
AG_FU1_y_p1 ~~ c(cov11)*AG_FU2_y_p1 + c(cov12)*AG_FU3_y_p1 
AG_FU2_y_p1 ~~ c(cov11)*AG_FU3_y_p1 

AG_IN_y_p2 ~~ c(cov21)*AG_FU1_y_p2 + c(cov22)*AG_FU2_y_p2 + c(cov23)*AG_FU3_y_p2 
AG_FU1_y_p2 ~~ c(cov21)*AG_FU2_y_p2 + c(cov22)*AG_FU3_y_p2 
AG_FU2_y_p2 ~~ c(cov21)*AG_FU3_y_p2 

AG_IN_y_p3 ~~ c(cov31)*AG_FU1_y_p3 + c(cov32)*AG_FU2_y_p3 + c(cov33)*AG_FU3_y_p3 
AG_FU1_y_p3 ~~ c(cov31)*AG_FU2_y_p3 + c(cov32)*AG_FU3_y_p3 
AG_FU2_y_p3 ~~ c(cov31)*AG_FU3_y_p3 

AG_IN_o_p1 ~~ c(cov11)*AG_FU1_o_p1 + c(cov12)*AG_FU2_o_p1 + c(cov13)*AG_FU3_o_p1 
AG_FU1_o_p1 ~~ c(cov11)*AG_FU2_o_p1 + c(cov12)*AG_FU3_o_p1 
AG_FU2_o_p1 ~~ c(cov11)*AG_FU3_o_p1 

AG_IN_o_p2 ~~ c(cov21)*AG_FU1_o_p2 + c(cov22)*AG_FU2_o_p2 + c(cov23)*AG_FU3_o_p2 
AG_FU1_o_p2 ~~ c(cov21)*AG_FU2_o_p2 + c(cov22)*AG_FU3_o_p2 
AG_FU2_o_p2 ~~ c(cov21)*AG_FU3_o_p2 

AG_IN_o_p3 ~~ c(cov31)*AG_FU1_o_p3 + c(cov32)*AG_FU2_o_p3 + c(cov33)*AG_FU3_o_p3 
AG_FU1_o_p3 ~~ c(cov31)*AG_FU2_o_p3 + c(cov32)*AG_FU3_o_p3 
AG_FU2_o_p3 ~~ c(cov31)*AG_FU3_o_p3 
'
#ad_sibAG <- sem(sibAG, data = yo[which(yo$adopt == 1),], missing = "FIML")
#summary(ad_sibAG, fit.measures = T, standardized = T)

#bio_sibAG <- sem(sibAG, data = yo[which(yo$adopt == 0),], missing = "FIML")
#summary(bio_sibAG, fit.measures = T, standardized = T)

#g_sibAG <- sem(sibAG, data = yo, missing = "FIML", group = "adopt", em.iter.max = 20000)
#summary(g_sibAG, fit.measures = T, standardized = T, ci = T)

sibAG <- sem(sibAG, data = yo, missing = "FIML")

fitMeasures(sibAG, c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"), 
            output = "text")

parameterEstimates(sibAG, standardized = T) %>% 
  filter(op == "~1" & lhs %in% c("iy", "io", "sy", "so")) %>% 
  select("Parameters" = lhs, B = est, SE = se, Z = z, "p-value" = pvalue, "CI lower" = ci.lower, "CI upper" = ci.upper) %>% 
  knitr::kable(caption = "Intercept and Slope") %>% 
  kable_styling()

parameterEstimates(sibAG, standardized = T) %>% 
  filter(op == "~~" & lhs %in% c("sy", "iy") & rhs %in% c("so", "io") & est > 0) %>% 
  select("Younger sibling" = lhs, "Older sibling" = rhs, B = est, SE = se, Z = z, "p-value" = pvalue, "CI lower" = ci.lower, "CI upper" = ci.upper) %>% 
  knitr::kable(caption = "Correlated intercepts and slopes") %>% 
  kable_styling()

semPaths(sibAG, what = "col", whatLabels = "est", intercepts = T)
```


### Control CON

```{r adopt-CON, warning = F, cache = T}
sibCON <- '
#younger sibling
CON0y =~   CON_IN_y_p1  + p2 * CON_IN_y_p2   + p3 * CON_IN_y_p3
CON3y =~   CON_FU1_y_p1 + p2 * CON_FU1_y_p2  + p3 * CON_FU1_y_p3
CON7y =~   CON_FU2_y_p1 + p2 * CON_FU2_y_p2  + p3 * CON_FU2_y_p3
CON16y =~  CON_FU3_y_p1 + p2 * CON_FU3_y_p2  + p3 * CON_FU3_y_p3
                   
#older sibling
CON0o =~   CON_IN_o_p1  + p2 * CON_IN_o_p2   + p3 * CON_IN_o_p3
CON3o =~   CON_FU1_o_p1 + p2 * CON_FU1_o_p2  + p3 * CON_FU1_o_p3
CON7o =~   CON_FU2_o_p1 + p2 * CON_FU2_o_p2  + p3 * CON_FU2_o_p3
CON16o =~  CON_FU3_o_p1 + p2 * CON_FU3_o_p2  + p3 * CON_FU3_o_p3

#latent intercept and slopes
iy =~ 1*CON0y + 1*CON3y + 1*CON7y + 1*CON16y
io =~ 1*CON0o + 1*CON3o + 1*CON7o + 1*CON16o
sy =~ 0*CON0y + 3.3*CON3y + 7.6*CON7y + 17*CON16y
so =~ 0*CON0o + 3.3*CON3o + 7.6*CON7o + 17*CON16o
iy ~ 1
io ~ 1
sy ~ 1
so ~ 1

#fix zero latent intercepts
CON0y ~ 0*1
CON3y ~ 0*1
CON7y ~ 0*1
CON16y ~ 0*1
CON0o ~ 0*1
CON3o ~ 0*1
CON7o ~ 0*1
CON16o ~ 0*1

#fix zero/equal item intercepts
CON_IN_y_p1  ~ 0*1
CON_FU1_y_p1 ~ 0*1
CON_FU2_y_p1 ~ 0*1
CON_FU3_y_p1 ~ 0*1
CON_IN_o_p1  ~ 0*1
CON_FU1_o_p1 ~ 0*1
CON_FU2_o_p1 ~ 0*1
CON_FU3_o_p1 ~ 0*1

CON_IN_y_p2  ~ par2*1
CON_FU1_y_p2 ~ par2*1
CON_FU2_y_p2 ~ par2*1
CON_FU3_y_p2 ~ par2*1
CON_IN_o_p2  ~ par2*1
CON_FU1_o_p2 ~ par2*1
CON_FU2_o_p2 ~ par2*1
CON_FU3_o_p2 ~ par2*1

CON_IN_y_p3  ~ par3*1
CON_FU1_y_p3 ~ par3*1
CON_FU2_y_p3 ~ par3*1
CON_FU3_y_p3 ~ par3*1
CON_IN_o_p3  ~ par3*1
CON_FU1_o_p3 ~ par3*1
CON_FU2_o_p3 ~ par3*1
CON_FU3_o_p3 ~ par3*1

#residual covariances
CON_IN_y_p1 ~~ c(cov11)*CON_FU1_y_p1 + c(cov12)*CON_FU2_y_p1 + c(cov13)*CON_FU3_y_p1 
CON_FU1_y_p1 ~~ c(cov11)*CON_FU2_y_p1 + c(cov12)*CON_FU3_y_p1 
CON_FU2_y_p1 ~~ c(cov11)*CON_FU3_y_p1 

CON_IN_y_p2 ~~ c(cov21)*CON_FU1_y_p2 + c(cov22)*CON_FU2_y_p2 + c(cov23)*CON_FU3_y_p2 
CON_FU1_y_p2 ~~ c(cov21)*CON_FU2_y_p2 + c(cov22)*CON_FU3_y_p2 
CON_FU2_y_p2 ~~ c(cov21)*CON_FU3_y_p2 

CON_IN_y_p3 ~~ c(cov31)*CON_FU1_y_p3 + c(cov32)*CON_FU2_y_p3 + c(cov33)*CON_FU3_y_p3 
CON_FU1_y_p3 ~~ c(cov31)*CON_FU2_y_p3 + c(cov32)*CON_FU3_y_p3 
CON_FU2_y_p3 ~~ c(cov31)*CON_FU3_y_p3 

CON_IN_o_p1 ~~ c(cov11)*CON_FU1_o_p1 + c(cov12)*CON_FU2_o_p1 + c(cov13)*CON_FU3_o_p1 
CON_FU1_o_p1 ~~ c(cov11)*CON_FU2_o_p1 + c(cov12)*CON_FU3_o_p1 
CON_FU2_o_p1 ~~ c(cov11)*CON_FU3_o_p1 

CON_IN_o_p2 ~~ c(cov21)*CON_FU1_o_p2 + c(cov22)*CON_FU2_o_p2 + c(cov23)*CON_FU3_o_p2 
CON_FU1_o_p2 ~~ c(cov21)*CON_FU2_o_p2 + c(cov22)*CON_FU3_o_p2 
CON_FU2_o_p2 ~~ c(cov21)*CON_FU3_o_p2 

CON_IN_o_p3 ~~ c(cov31)*CON_FU1_o_p3 + c(cov32)*CON_FU2_o_p3 + c(cov33)*CON_FU3_o_p3 
CON_FU1_o_p3 ~~ c(cov31)*CON_FU2_o_p3 + c(cov32)*CON_FU3_o_p3 
CON_FU2_o_p3 ~~ c(cov31)*CON_FU3_o_p3 
'

sibCON <- sem(sibCON, data = yo, missing = "FIML")

fitMeasures(sibCON, c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"), 
            output = "text")

parameterEstimates(sibCON, standardized = T) %>% 
  filter(op == "~1" & lhs %in% c("iy", "io", "sy", "so")) %>% 
  select("Parameters" = lhs, B = est, SE = se, Z = z, "p-value" = pvalue, "CI lower" = ci.lower, "CI upper" = ci.upper) %>% 
  knitr::kable(caption = "Intercept and Slope") %>% 
  kable_styling()

parameterEstimates(sibCON, standardized = T) %>% 
  filter(op == "~~" & lhs %in% c("sy", "iy") & rhs %in% c("so", "io") & z < 1.324) %>% 
  select("Younger sibling" = lhs, "Older sibling" = rhs, B = est, SE = se, Z = z, "p-value" = pvalue, "CI lower" = ci.lower, "CI upper" = ci.upper) %>% 
  knitr::kable(caption = "Correlated intercepts and slopes") %>% 
  kable_styling()

semPaths(sibCON, what = "col", whatLabels = "est", intercepts = T)
```


### Harm Avoidance HA

```{r adopt-HA, warning = F, cache = T}
sibHA <- '
#younger sibling
HA0y =~   HA_IN_y_p1  + p2 * HA_IN_y_p2   + p3 * HA_IN_y_p3
HA3y =~   HA_FU1_y_p1 + p2 * HA_FU1_y_p2  + p3 * HA_FU1_y_p3
HA7y =~   HA_FU2_y_p1 + p2 * HA_FU2_y_p2  + p3 * HA_FU2_y_p3
HA16y =~  HA_FU3_y_p1 + p2 * HA_FU3_y_p2  + p3 * HA_FU3_y_p3
                   
#older sibling
HA0o =~   HA_IN_o_p1  + p2 * HA_IN_o_p2   + p3 * HA_IN_o_p3
HA3o =~   HA_FU1_o_p1 + p2 * HA_FU1_o_p2  + p3 * HA_FU1_o_p3
HA7o =~   HA_FU2_o_p1 + p2 * HA_FU2_o_p2  + p3 * HA_FU2_o_p3
HA16o =~  HA_FU3_o_p1 + p2 * HA_FU3_o_p2  + p3 * HA_FU3_o_p3

#latent intercept and slopes
iy =~ 1*HA0y + 1*HA3y + 1*HA7y + 1*HA16y
io =~ 1*HA0o + 1*HA3o + 1*HA7o + 1*HA16o
sy =~ 0*HA0y + 3.3*HA3y + 7.6*HA7y + 17*HA16y
so =~ 0*HA0o + 3.3*HA3o + 7.6*HA7o + 17*HA16o
iy ~ 1
io ~ 1
sy ~ 1
so ~ 1

#fix zero latent intercepts
HA0y ~ 0*1
HA3y ~ 0*1
HA7y ~ 0*1
HA16y ~ 0*1
HA0o ~ 0*1
HA3o ~ 0*1
HA7o ~ 0*1
HA16o ~ 0*1

#fix zero/equal item intercepts
HA_IN_y_p1  ~ 0*1
HA_FU1_y_p1 ~ 0*1
HA_FU2_y_p1 ~ 0*1
HA_FU3_y_p1 ~ 0*1
HA_IN_o_p1  ~ 0*1
HA_FU1_o_p1 ~ 0*1
HA_FU2_o_p1 ~ 0*1
HA_FU3_o_p1 ~ 0*1

HA_IN_y_p2  ~ par2*1
HA_FU1_y_p2 ~ par2*1
HA_FU2_y_p2 ~ par2*1
HA_FU3_y_p2 ~ par2*1
HA_IN_o_p2  ~ par2*1
HA_FU1_o_p2 ~ par2*1
HA_FU2_o_p2 ~ par2*1
HA_FU3_o_p2 ~ par2*1

HA_IN_y_p3  ~ par3*1
HA_FU1_y_p3 ~ par3*1
HA_FU2_y_p3 ~ par3*1
HA_FU3_y_p3 ~ par3*1
HA_IN_o_p3  ~ par3*1
HA_FU1_o_p3 ~ par3*1
HA_FU2_o_p3 ~ par3*1
HA_FU3_o_p3 ~ par3*1

#residual covariances
HA_IN_y_p1 ~~ c(cov11)*HA_FU1_y_p1 + c(cov12)*HA_FU2_y_p1 + c(cov13)*HA_FU3_y_p1 
HA_FU1_y_p1 ~~ c(cov11)*HA_FU2_y_p1 + c(cov12)*HA_FU3_y_p1 
HA_FU2_y_p1 ~~ c(cov11)*HA_FU3_y_p1 

HA_IN_y_p2 ~~ c(cov21)*HA_FU1_y_p2 + c(cov22)*HA_FU2_y_p2 + c(cov23)*HA_FU3_y_p2 
HA_FU1_y_p2 ~~ c(cov21)*HA_FU2_y_p2 + c(cov22)*HA_FU3_y_p2 
HA_FU2_y_p2 ~~ c(cov21)*HA_FU3_y_p2 

HA_IN_y_p3 ~~ c(cov31)*HA_FU1_y_p3 + c(cov32)*HA_FU2_y_p3 + c(cov33)*HA_FU3_y_p3 
HA_FU1_y_p3 ~~ c(cov31)*HA_FU2_y_p3 + c(cov32)*HA_FU3_y_p3 
HA_FU2_y_p3 ~~ c(cov31)*HA_FU3_y_p3 

HA_IN_o_p1 ~~ c(cov11)*HA_FU1_o_p1 + c(cov12)*HA_FU2_o_p1 + c(cov13)*HA_FU3_o_p1 
HA_FU1_o_p1 ~~ c(cov11)*HA_FU2_o_p1 + c(cov12)*HA_FU3_o_p1 
HA_FU2_o_p1 ~~ c(cov11)*HA_FU3_o_p1 

HA_IN_o_p2 ~~ c(cov21)*HA_FU1_o_p2 + c(cov22)*HA_FU2_o_p2 + c(cov23)*HA_FU3_o_p2 
HA_FU1_o_p2 ~~ c(cov21)*HA_FU2_o_p2 + c(cov22)*HA_FU3_o_p2 
HA_FU2_o_p2 ~~ c(cov21)*HA_FU3_o_p2 

HA_IN_o_p3 ~~ c(cov31)*HA_FU1_o_p3 + c(cov32)*HA_FU2_o_p3 + c(cov33)*HA_FU3_o_p3 
HA_FU1_o_p3 ~~ c(cov31)*HA_FU2_o_p3 + c(cov32)*HA_FU3_o_p3 
HA_FU2_o_p3 ~~ c(cov31)*HA_FU3_o_p3 
'

sibHA <- sem(sibHA, data = yo, missing = "FIML")

fitMeasures(sibHA, c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"), 
            output = "text")

parameterEstimates(sibHA, standardized = T) %>% 
  filter(op == "~1" & lhs %in% c("iy", "io", "sy", "so")) %>% 
  select("Parameters" = lhs, B = est, SE = se, Z = z, "p-value" = pvalue, "CI lower" = ci.lower, "CI upper" = ci.upper) %>% 
  knitr::kable(caption = "Intercept and Slope") %>% 
  kable_styling()

parameterEstimates(sibHA, standardized = T) %>% 
  filter(op == "~~" & lhs %in% c("sy", "iy") & rhs %in% c("so", "io") & est > 0) %>% 
  select("Younger sibling" = lhs, "Older sibling" = rhs, B = est, SE = se, Z = z, "p-value" = pvalue, "CI lower" = ci.lower, "CI upper" = ci.upper) %>% 
  knitr::kable(caption = "Correlated intercepts and slopes") %>% 
  kable_styling()

semPaths(sibHA, what = "col", whatLabels = "est", intercepts = T)
```